A Handbook for Tenured Track Faculty in Computer Science in the US

### 2024-07-08 faculty.tex
updates

I also looked at CVs of other people in the dept. who recently got tenure. I also sent my CV to  my mentors and got many feedback. In fact, I also asked a Division Dean in the College who is in charge of P\&T procedure to review my CV for formatting and other issues. In short, the CV is one of the easiest part of the tenure package, and so I want to make sure mine has no silly mistakes.

@@

Here is my \href{https://dynaroars.github.io/people/nguyenthanhvuh/files/pt/statement-annotated.pdf}{tenure application statement}, annotated with comments. At GMU, the statement is limited to 8 pages.  My statement talks about my achievements and plans in (i) research and funding, (ii) advising and mentoring, (iii) teaching, and (iv) services. This organization is somewhat standard, and the main unique part of my statement is the first page where I highlight and summarize my achievements (e.g., papers, fundings, students, etc). I got these information and numbers from my CV.

While preparing the statement was the most time-consuming part of my tenure package, it wasn't too bad and may be just took a couple of days.  I was able to reuse lots of text and contents from my job application statements (when I applied to GMU in 2021). I also got statements from colleagues at GMU who recently got tenure and a couple of people at other universities to get some ideas on what to write.  I did not send my statement for feedback because I felt that it was good enough and I didn't want to bother people (reading an 8-page statement is a lot more work).

\subsection{Letters}
This is a huge part of your tenure package.  You will need letters from senior people in your area.  Someone who can vouch for you.
You will need 4--6 letters. However, you will not be the one asking for these letters, but your department will. You can suggest who to ask.

Typically, there are \emph{two lists} of letter writers: one consisting of people suggested by you and the other consisting of people suggested by the department (i.e., your dept will come up with a list of people they think are good letter writers for you). There can also be another list, the ``blacklist'' where you list people that you don't want to write letters for you.

\subsubsection{Personal Experience}
Originally I asked several senior people to write my letters. Then I found out that there is a rule at GMU that I cannot ask people whom have collaboration with me in the last 7 years!!! That invalidates most of my letter writers.  Moreover, GMU (and likely most other universities) would prefer writers not from my list (because people on my list are likely to be biased), and has a rule that at least 60\% of the letter writers must be from the department's list. Moreover, there's a good chance that a letter writer will decline the invitation, so there should be more names.

After talking to various people in my dept and outside, I decide to put just 2 names in my list (0 on blacklist) and let the department decides the rest. I also thought about leaving my list empty but then that seems to be too weird, and I don't want to give the impression that I don't have any friends in the world.
The reason is that I feel I don't have too many "enemies" out there that would write bad things about me.  Those that don't know me well will likely just decline the letter invitation. Turned out to be a good decision because in the end the department selected 1 of the 2 names on my list and 4 more names from the department's list. Moreover, it looks like no one declines so my letter writers are all set fairly quickly in the process.  All of my letters were good and give many positive comments about me as summarized in my evaluation letter (sent to me after the tenure decision).  I will never know who wrote the letters, but I am very grateful to them.

@@

\paragraph{Tenure Visits} Many faculty will visit other universities to give talks and meet people.  This is a good way to get your name out and to get letters (e.g., from the person that host you).  Physical visits were not possible during the pandemic, but virtual visits are possible and in fact easier. However,  I did not do any tenure visits or give talks in the year leading up to my tenure application. This was because I am not a travel person and also I felt that it would not help much in my case.

>>>>

I also looked at CVs of other people in the dept. who recently got tenure. I also sent my CV to  my mentors and got many feedback. In fact, I also asked a Division Dean in the College who is in charge of P\&T procedure to review my CV for formatting and other issues. In short, the CV is the easiest part of my tenure package, and so I want to make sure mine has no silly mistakes.

@@

You will need to prepare a statement for your tenure package.  Typically, the statement talks about your achievements and plans in (i) research and funding, (ii) advising and mentoring, (iii) teaching, and (iv) services. This order is somewhat standard, but you can change it if you want. In a sense, this statement is similar to your job application statement, but it is more focused on what you have done (and less on what you will do). However, unlike job application, tenure statement has a page limit (e.g., 8 pages at GMU).

Here is my \href{https://dynaroars.github.io/people/nguyenthanhvuh/files/pt/statement-annotated.pdf}{tenure application statement}, annotated with comments. The unique part of my statement is the first page where I highlight and summarize my achievements (e.g., papers, fundings, students, etc). I got these information and numbers from my CV.
While preparing the statement was the most time-consuming part of my tenure package, it wasn't too bad and just took a couple of days.  I was able to reuse lots of text and contents from my job application statements (when I applied to GMU in 2021). I also got statements from colleagues at GMU who recently got tenure and a couple of people at other universities to get some ideas on what to write.  I did not send my statement for feedback because I felt that it was good enough and I didn't want to bother people (reading an 8-page statement is a lot more work).

\subsection{Letters}
You will need letters from senior people in your area.  Someone who can vouch for you.
You will need 4--6 letters. However, you will not be the one asking for these letters, but your department will. You can suggest who to ask.  This process doesn't take much of your time (because you don't write the letters) but it might cause you a lot of stress (because you don't have full control over it).

Typically, your department will compile \emph{two lists} of letter writers: one consisting of people suggested by you and the other consisting of people suggested by the department (i.e., your dept will come up with a list of people they think are good letter writers for you). The people on the departmental list often "weight" more than those on your list.
There can also be another list, the ``blacklist'' where you list people that you don't want to write letters for you.

\begin{commentbox}[Vu:]
Originally I asked several senior people to write my letters. Then I found out that there is a rule at GMU that I cannot ask people whom have collaboration with me in the last 7 years!!! That invalidates most of my letter writers.  Moreover, GMU (and likely most other universities) would prefer writers not from my list (because people on my list are likely to be biased), and has a rule that at least 60\% of the letter writers must be from the department's list. There's also a good chance that a letter writer will decline the invitation, so there should be more names.
\end{commentbox}

In my case I put just 2 names in my list (0 on blacklist) and let the department decide the rest. The reason that I only put 2 names is that I feel I don't have too many "enemies" out there that would write bad things about me.  Those that don't know me well will likely just decline the letter invitation.  This also gives the department more choices on their list.
Note I also thought about leaving my list empty, but then that seems to be too weird, and I don't want to give the impression that I don't have any "friends" in academic.

The department selected 1 of the 2 names on my list and 4 more names from the department's list. Moreover, it appears that no one declines so my letter writers are all set fairly quickly in the process. All of my letters were very good and give many positive comments about me as summarized in my evaluation letter (sent to me after the tenure decision).  I will never know who wrote the letters, but I am very grateful to them.

\subsection{Tenure Talk}
After you submit your tenure package, you might be asked to give a tenure talk.  This is a talk where you summarize your research, teaching, and services, and future plans.  The talk is typically 1 hour long, and there will be a Q\&A session afterward.  The talk is open to the public.

@@

\paragraph{Tenure Visits} Many people will visit other universities to give talks and meet people.  This is a good way to get your name out and to get letters (e.g., from the person that host you).  Physical visits were not possible during the pandemic, but virtual visits are possible and in fact easier. However,  I did not do any tenure visits or give talks in the year leading up to my tenure application. This was because I am not a travel person and also I felt that it would not help much in my case.
